---
title: "Exercise Behavior and Prediction"
author: "Dahyr J. Vergara Suárez"
date: "Friday, September 19, 2014"
output: html_document
---


### Executive Summary



### Data Gathering & Exploration

```{r get-data, cache=FALSE, echo=FALSE}
library(ggplot2)
library(caret)
```


```{r get-data, cache=TRUE, echo=FALSE}
setInternet2(TRUE)
download.file(
	"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
	"pml-training.csv")
training <- read.csv("pml-training.csv")
download.file(
	"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
	"pml-testing.csv")
testing <- read.csv("pml-testing.csv")
setDim <- rbind(dim(testing), dim(training))
rownames(setDim) <- c("Testing Set", "Training Set")
colnames(setDim) <- c("# Obs.", "# Vars.")
setDim

# Starts on sunday & january
training$formatted_time <- as.POSIXlt(training$cvtd_timestamp, 
									  format = "%d/%m/%Y %H:%M")
training$weekday <- training$formatted_time$wday + 1
training$month <- training$formatted_time$mon + 1
training$hour <- training$formatted_time$hour
training$new_window <- as.numeric(training$new_window == "yes")


testing$formatted_time <- as.POSIXlt(testing$cvtd_timestamp, 
									  format = "%d/%m/%Y %H:%M")
testing$weekday <- testing$formatted_time$wday + 1
testing$month <- testing$formatted_time$mon + 1
testing$hour <- testing$formatted_time$hour
testing$new_window <- as.numeric(testing$new_window == "yes")

# Subsetting to exclude some variables that we know won't affect the outcome
training <- subset(training, select = -c(1, user_name, raw_timestamp_part_1, 
				 	raw_timestamp_part_2, cvtd_timestamp, formatted_time))
testing <- subset(testing, select = -c(1, user_name, raw_timestamp_part_1, 
				 	raw_timestamp_part_2, cvtd_timestamp, formatted_time))

# Correcting zero length variables and NA's only for the training set
for(i in 3:ncol(training)) {
	if(class(training[, i]) == "factor" && names(training)[i] != "classe") {
		trainNA <- training[, i] == ""
		trainInf <- training[, i] == "#DIV/0!"
		training[trainNA, i] <- NA
		training[trainInf, i] <- Inf
		training[, i] <- as.numeric(as.character(training[, i]))
	}
}


```


Now let's remove all those variables that have near zero variance and then use principal component analysis (PCA) with thresholds of 70, 80 and 90% of variance to find a good number of variables to use in our predictions. With 70% variance we can use 14, for 80% 21 and to obtain variance of 90% 34 variables would be needed. With this in mind the decision is to use 14 variables


```{r pca, fig.width=10, fig.height=10}
## Variance analysis
# Subsetting again to exclude those variables with near zero variance
nzv <- nearZeroVar(training, saveMetrics = T)
training <- training[, rownames(nzv)[!nzv$nzv]]

pp70 <- preProcess(training[, -119], method = "pca", thresh = 0.7)
pp80 <- preProcess(training[, -119], method = "pca", thresh = 0.8)
pp90 <- preProcess(training[, -119], method = "pca", thresh = 0.9)

pcaComps <- as.data.frame(c(pp70$numComp, pp80$numComp, pp90$numComp))
names(pcaComps) <- c("# Components")
rownames(pcaComps) <- c("70% Variation", "80% Variation", "90% Variation")
pcaComps

predpca70 <- predict(pp70, training[, -119])
qplot(predpca70[, 1], predpca70[, 2], col = training$classe)
#confusionMatrix(training$classe, predict(pp70, predpca70))

rf <- train(classe ~ ., data = training, method = "rf", 
			prox = T, importance = T)

rfx3 <- train(classe ~ ., data = training, method = "rf", 
			  ntree = 1500, prox = T)

rfpca <- train(classe ~ ., data = training, method = "rf", preProcess = "pca", 
			   prox = T, thresh = 0.9)

topImportanceRf <- importance(rf$finalModel)[, c(6, 7)]

head(topImportanceRf[order(-topImportanceRf[, 2]), ], 5)

qplot(avg_roll_dumbbell, avg_pitch_belt, color = classe, data = training)
#library(rattle)
#fancyRpartPlot(getTree(rf$finalModel, k = 1))
```

Resampling is done within the implementation of the random forests algorithm (by default 500 trees)


> rfx3 <- train(classe ~ ., data = training, method = "rf", 
+ 			  ntree = 1500, prox = T)
> rfx3
Random Forest 

19622 samples
  121 predictors
    5 classes: 'A', 'B', 'C', 'D', 'E' 

No pre-processing
Resampling: Bootstrapped (25 reps) 

Summary of sample sizes: 217, 217, 217, 217, 217, 217, ... 

Resampling results across tuning parameters:

  mtry  Accuracy  Kappa  Accuracy SD  Kappa SD
  2     0.75      0.684  0.0578       0.0711  
  61    0.741     0.673  0.056        0.0688  
  121   0.733     0.663  0.0558       0.0687  

Accuracy was used to select the optimal model using  the largest value.
The final value used for the model was mtry = 2. 


```{r healthdamage, fig.width=10, fig.height=10}
qplot(gyros_belt_x, classe, data = training)
```


### Model Building

- Model description


- Cross validation



```{r build-model, cache=TRUE, echo=FALSE}

```


- Expected sample error




### Model Testing

- 20 test cases


- Misfits


- Actual error


```{r, echo=FALSE}

```



